Mergeable Types

### 2017-02-25 paper/mergeable-types.tex
edits

  \item There exists a function $M: (t \rightarrow t)^*\!\times\!(t
  \rightarrow t)^* \;\rightarrow\; (t \rightarrow t)\!\times\!(t
  \rightarrow t)$ that accepts a pair of minimal sequences of
  primitive morphisms, $P^*$ and $Q^*$, and returns a pair of morphisms,
  $P'$ and $Q'$, such that for any object $A:t$, $(Q' \circ
  \denot{P^*})(A) = (P' \circ \denot{Q^*})(A)$.  
\end{itemize}
Commutativity diagram of Fig.~\ref{fig:pushouts-2} visualizes this
definition. 
\end{definition}

The definition asserts the existence of two functions for a mergeable
type $t$. First, the function $W$, which computes the \emph{edit
distance} between an object $A:t$ and its semantic successor $B:t$.
The sequence of primitive morphisms it returns ($P^*$) is called an
\emph{edit script}. Replaying the edit script on $A$ produces $B$
(i.e., $\denot{P^*}(A) = B$). Second, the definition asserts the
existence of a function $M$, which performs an \emph{operational
transformation} of the primitive morphisms in an edit script. An
\begin{wrapfigure}{l}{.4\textwidth}
\centering
\includegraphics[scale=0.4]{Figures/string-eg}
\caption{Merging strings with edit scripts and operational
transformations}
\label{fig:string-eg}
\end{wrapfigure}
operational transformation of a primitive morphism $P$ w.r.t a
morphism $Q$ is another primitive morphism $P'$ such that for any
object $A:t$, $P'$ has the same \emph{effect} on $Q(A)$ as $P$ has on
$A$. We will now consider examples of mergeable types that reify the
notions of edit scripts and operational transformations.
\paragraph{Strings} Strings are mergeable. Given a string ($s$) and
its semantic successor ($s'$), an edit script ($P^*$ or $Q^*$) is the
shortest sequence of \C{insert}, \C{delete}, and \C{substitute}
operations (a string's primitive morphisms), whose composition takes $s$
to $s'$. The length of the sequence is in fact the Levenshtein
distance~\cite{levenstein} between $s$ and $s'$, which can be computed
in polynomial time~\cite{wagner-fischer}.
A slight modification of Wagner-Fischer also lets us compute the edit
script. Fig.~\ref{fig:string-eg} illustrates.

>>>>

  \item There exists a function $M: (t \rightarrow t)^*\!\times\!(t
  \rightarrow t)^* \;\rightarrow\; (t \rightarrow t)\!\times\!(t
  \rightarrow t)$ that accepts a pair of minimal sequences of
  primitive morphisms, $P^*$ and $Q^*$, and returns a pair of morphisms,
  $P'$ and $Q'$, such that for any object $A:t$, $(Q' \circ
  \denot{P^*})(A) = (P' \circ \denot{Q^*})(A)$.  
\end{itemize}
The Commutativity diagram of Fig.~\ref{fig:pushouts-2} visualizes this
definition. 
\end{definition}

The definition asserts the existence of two functions for a mergeable
type $t$. First, the function $W$, which computes the \emph{edit
distance} between an object $A:t$ and its semantic successor $B:t$.
The sequence of primitive morphisms it returns ($P^*$) is called an
\emph{edit script}. Replaying the edit script on $A$ produces $B$
(i.e., $\denot{P^*}(A) = B$). Second, the definition asserts the
existence of a function $M$, which performs an \emph{operational
transformation} of the primitive morphisms in an edit script. An
operational transformation of a primitive morphism $P$ w.r.t a
morphism $Q$ is another primitive morphism $P'$ such that for any
object $A:t$, $P'$ has the same \emph{effect} on $Q(A)$ as $P$ has on
$A$. We will now consider examples of mergeable types that reify the
notions of edit scripts and operational transformations.
\begin{wrapfigure}{l}{.4\textwidth}
\centering
\includegraphics[scale=0.4]{Figures/string-eg}
\caption{Merging strings with edit scripts and operational
transformations}
\label{fig:string-eg}
\end{wrapfigure}
\paragraph{Strings} Strings are mergeable. Given a string ($s$) and
its semantic successor ($s'$), an edit script ($P^*$ or $Q^*$) is the
shortest sequence of \C{insert}, \C{delete}, and \C{substitute}
operations (a string's primitive morphisms), whose composition takes $s$
to $s'$. The length of the sequence is in fact the Levenshtein
distance~\cite{levenstein} between $s$ and $s'$, which can be computed
in polynomial time~\cite{wagner-fischer}.
A slight modification of Wagner-Fischer also lets us compute the edit
script. Fig.~\ref{fig:string-eg} illustrates.
